# Workshop 1: random intercept models met INLA

Zie <https://inbo.github.io/tutorials/tutorials/r_inla/>

We beperken de data voorlopig tot de LG3-respons.

```{r}
lg3data <- 
    XG3data %>% 
    filter(Variable == "LG3") %>% 
    rename(Response = Value) %>% 
    select(-Variable)
lg3data %>% str
```


## Challenge 1

### fit fixed effect model

```{r}
I1 <- inla(Response ~ Type + Soiltexture + HYear,
           data = lg3data,
           control.compute = list(dic = TRUE, waic = TRUE),
           family = "gaussian") # gaussian is the default!
summary(I1)
```


### extract $WAIC$ from the model

```{r}
I1 %>% names
```

```{r}
I1$waic %>% str
```

```{r}
I1$waic$waic
```


### display fixed effect parameters in a table

```{r}
I1$summary.fixed %>% 
    rownames_to_column(var = "Fixed_effect") %>% 
    select(Fixed_effect, mean, `0.025quant`, `0.975quant`) %>% 
    kable
```




## Challenge 2

### what is the default prior for a fixed effect (`?control.fixed`)
### use a custom prior for a fixed effect (`?inla`)
### specify two linear combinations [r-inla.org, FAQ 17](http://www.r-inla.org/faq#TOC-I-have-some-linear-combinations-of-the-nodes-in-the-latent-field-that-I-want-to-compute-the-posterior-marginal-of-is-that-possible-)


## Challenge 3

### fit a model with one or more random intercepts (`model = 'iid'`)
### what is the default prior for `'iid'` (`inla.doc('iid')`)
### calculate $\sigma$ for the random intercept
### display the BLUP the random intercept

## Challenge 4

### Think about the relevant magnitude of $\sigma$ for your random effect
### Use a custom "pc.prec" prior with that $\sigma$ (`inla.doc("pc.prec")`)


## Challenge 5

### pick a relevant variable for an 'rw1' model
### ponder on a relevant $\sigma$ for that model
### fit model with 'rw1' component and `pc.prec` prior
