# Workshop 1: random intercept models met INLA

Zie <https://inbo.github.io/tutorials/tutorials/r_inla/>

We beperken de data voorlopig tot de LG3-respons.

```{r}
lg3data <- 
    XG3data %>% 
    filter(Variable == "LG3") %>% 
    rename(Response = Value) %>% 
    select(-Variable)
lg3data %>% str
```


## Challenge 1

### fit fixed effect model

```{r}
I1 <- inla(Response ~ Type + Soiltexture + HYear,
           data = lg3data,
           control.compute = list(dic = TRUE, waic = TRUE),
           family = "gaussian") # gaussian is the default!
summary(I1)
```


### extract $WAIC$ from the model

```{r}
I1 %>% names
```

```{r}
I1$waic %>% str
```

```{r}
I1$waic$waic
```


### display fixed effect parameters in a table

```{r}
I1$summary.fixed %>% 
    rownames_to_column(var = "Fixed_effect") %>% 
    select(Fixed_effect, mean, `0.025quant`, `0.975quant`) %>% 
    kable
```

Citaat uit script `Ex719C_Otolith_V4.R` van A. Zuur:

```
# To make the numerical estimation process 
# easier, we standardize all continous 
# covariates.
```

Laten we dus het volgende proberen en het resultaat vergelijken:

```{r}
MyStd <- function(x) { (x - mean(x)) / sd(x)} # van A. Zuur
lg3data <- 
    lg3data %>% 
    mutate(HYear_std = MyStd(HYear))
```

```{r}
inla(Response ~ Type + Soiltexture + HYear_std,
           data = lg3data,
           control.compute = list(dic = TRUE, waic = TRUE),
           family = "gaussian") %>% 
    .$summary.fixed %>% 
    rownames_to_column(var = "Fixed_effect") %>% 
    select(Fixed_effect, mean, `0.025quant`, `0.975quant`) %>% 
    kable
```

De coëfficiënt van HYear is uiteraard veranderd!
Ook het intercept is veranderd.
Overige parameters zijn niet duidelijk anders.

De snelheid is nauwelijks verbeterd.
Dit heeft mogelijks te maken met het feit dat `HYear` reeds relatief lage waarden had (variërend van 0 tot 27).

```{r}
system.time(inla(Response ~ Type + Soiltexture + HYear,
           data = lg3data,
           control.compute = list(dic = TRUE, waic = TRUE),
           family = "gaussian"))
```

```{r}
system.time(inla(Response ~ Type + Soiltexture + HYear_std,
           data = lg3data,
           control.compute = list(dic = TRUE, waic = TRUE),
           family = "gaussian"))
```

## Challenge 2

### what is the default prior for a fixed effect (`?control.fixed`)
### use a custom prior for a fixed effect (`?inla`)
### specify two linear combinations [r-inla.org, FAQ 17](http://www.r-inla.org/faq#TOC-I-have-some-linear-combinations-of-the-nodes-in-the-latent-field-that-I-want-to-compute-the-posterior-marginal-of-is-that-possible-)


## Challenge 3

### fit a model with one or more random intercepts (`model = 'iid'`)
### what is the default prior for `'iid'` (`inla.doc('iid')`)
### calculate $\sigma$ for the random intercept
### display the BLUP the random intercept

## Challenge 4

### Think about the relevant magnitude of $\sigma$ for your random effect
### Use a custom "pc.prec" prior with that $\sigma$ (`inla.doc("pc.prec")`)


## Challenge 5

### pick a relevant variable for an 'rw1' model
### ponder on a relevant $\sigma$ for that model
### fit model with 'rw1' component and `pc.prec` prior
